Bounds on Coding Theory\\from Algebraic Geometry

### 2021-05-06 draft.tex
2021-05-06 Morning Spark

\usepackage[english]{babel}
\usepackage{amsmath, amssymb, amsfonts, amsthm, graphicx}
\usepackage[mathscr]{eucal}

@@

\section{Singleton bound and a promising example} \label{s:singleton}
\begin{thm}[Singleton Bound]\label{thm:singleton_bound}
If $C$ is a code with parameters $n, k, d$, then 
	\begin{equation} % IF SPACE NEEDED, TRANSFORM THESE EQUATIONS INTO INLINE
		k + d \le n + 1,
	\end{equation} 
	or equivalently, 
	\begin{equation}
		R + \delta \le 1 + 1/n.
	\end{equation}
\end{thm}

@@

\begin{proof}
	$RS_q(n, k)$ is a subset of $\mathbb{F}_q^n$, thus it has length $n$.  
	Note that $ L_k$ is a vector space over $\mathbb{F}_q$.
	Note that $\{1, x, x^2, \dots, x^{k-1}\}$ is a choice of basis for this vector space, thus it has dimension $k$.
	Consider the map $\phi:  L_k \to \mathbb{F}_q^n$ given by 
	\begin{equation}
		f \mapsto (f(\alpha_1), f(\alpha_2), \dots, f(\alpha_n)).
	\end{equation}

	Note that the map $\phi$ is a linear transformation.
	Thus, its image $\Ima \phi = RS_q(n, k)$ is also a vector space.  
	Additionally, if $\phi(f) = \phi(g)$, then $f - g$ has at least $n$ roots, but has degree less than $n$; thus $f - g$ is the zero polynomial, which implies $f = g$.
	Therefore, $\phi$ is also injective.
	This implies that the dimension of the domain $ L_k$ is the same as the dimension of the image $RS_q(n, k)$, i.e.,  $\dimension RS_q(n, k) = k$.

	Finally, consider distinct $f, g \in  L_k$ and define $d = \dist(\phi(f), \phi(g))$, $f - g$ has at least $n - d$ roots.
	Furthermore, $f - g$ is a non-zero polynomial with degree less than $k$, thus has at most $k - 1$ roots.
	Then,  $ k - 1 \ge \#\ \text{roots} \ge n - d$.
	If we choose $f, g$ such that $d$ is the minimal distance, we get $k + d_{\mathrm{min}} \ge n + 1$, which together with \nameref{thm:singleton_bound} implies $k + d_\text{min} = n + 1$.
\end{proof}

@@

\section{Curves, Rational Functions and Divisors}

@@

\begin{defn}[Rational function]
	A rational function $f$ over $K$ is the ratio $g/h$ of polynomials $g, h \in K[x, y]$.
	A homogeneous rational function $F$ over a $K$ is the ratio $G/H$ of homogeneous polynomials $G, H \in K[x, y]$.

	After cancelling common roots of $g, h$, the roots of $g$ are called \emph{zeros} of $f$ and the roots of $h$ are called the \emph{poles} of $f$.  
	We say $f$ has order $n$ at $P$ if $P$ is a zero of multiplicity $n$; order $-n$ if $P$ is a pole with multiplicity $n$; and order $0$, otherwise.
\end{defn}

@@

\begin{defn}[Riemann--Roch space, \cite{LS87}]
	If $D = \sum n_p P$, then define the \emph{Riemann--Roch space $\mathcal L(D)$} as the vector space formed by all homogeneous rational functions $f$ such that the order of $f$ at each point $P$ of $\mathfrak C$ is greater or equal to $-n_P$ (plus the zero polynomial).
\end{defn}

@@

\begin{proof}
	\todo{Prove this Proposition.}
\end{proof}

For our study, an important theorem from Algebraic Geometry that will be useful in Section \ref{s:grs} when creating new types of codes is the following one:
\begin{thm}[Riemman--Roch Theorem, \cite{Wal00}]
	Let $\mathfrak C$ be a nonsingular projective curve of genus%
	\footnote{For the reader that is not familiar with genus, it is enough to know that it is an integer that can be calculated for any given curve.}
	$g$ defined over the field $\mathbb{F}_q$ and let $D$ be a divisor on $\mathfrak C$.
	Then
	\begin{equation}
		\dim \mathcal L(D) \ge \deg D + 1 - g,
	\end{equation}
	with equality holding if $\deg D > 2g - 2$.
\end{thm}

@@

\begin{prop}
	The sets $L_k$ and $\mathcal L(D)$ are mapped with a bijection (homogenization) $\phi: f(x) \mapsto Y^{\deg f} f(X/Y)$.
\end{prop}
\begin{proof}
	\todo{Write proof.}
\end{proof}

@@

\begin{defn}[Algebraic geometric codes, \cite{LS87, Wal00}] %Copied from Wal00
	Let $\mathfrak C$ be an irreducible nonsingular%
	\footnote{These adjectives just make sure $\mathfrak C$ does not have some weird behavior, e.g., $\mathfrak C$ has no double root. The reader should not worry much about them.}
%%	\todo{Provide an explanation for why the `irreducible nonsingular' is there, but not a thing that the reader should worry. I think the key is some sort of multiplicity.}
	projective plane curve, let $D$ be a divisor on $\mathfrak C$ and let $\mathcal P = \{P_1, P_2, \dots, P_n\}$ be a set of rational points on $\mathfrak C$.
	Assume $\mathcal P$ and $\Supp D$ have no points in common, thus, no $P_i$ can be a pole of any $f \in \mathcal L(D)$, and $f(P_i)$ is well-defined for any $f \in \mathcal L(D)$ and any $P_i \in \mathcal P$.
	Then, the \emph{algebraic geometric code} associated to $\mathfrak C, \mathcal P$ and $D$ is
	\begin{equation}
		C(\mathfrak C, \mathcal P, D) := \{(f(P_1), \dots, f(P_n)) \mid f \in L(D)\} \subset \mathbb{F}_q^n.
	\end{equation}
\end{defn}

@@

	Futhermore, we shall prove that $\psi$ is injective.
	If $\psi(f) = \psi(g)$, then $\psi(f - g) = \vec 0$.
	Let $h := f-g$.
	We have $h(P_1) = h(P_2) = \dots = h(P_n) = 0$, so every $P_i$ is a zero of $h$, and finally $a_{P_i} \ge 1$, for all $1 \le i \le n$.
	Therefore, $h \in \mathcal L(D - P_1 - \cdots - P_n)$.
	Since $\deg D < n$, we have $\deg(D - P_1 - \cdots - P_n) < 0$ and, consequenly, $h$ must be the zero polynomial by Proposition \ref{prop:degDlt0}.
	Therefore, $f = g$, finishing the proof that $\psi$ is injective.
	Since $\psi$ is a injective linear transformation, the dimension of the domain and of the image are the same.
	Therefore, the dimension of $C(\mathfrak C, \mathcal P, D)$ is $k = \dim \mathcal L(D)$, which evaluates to $\deg D + 1 - g$ by the Riemman-Roch Theorem.

>>>>

\usepackage[english]{babel}
\usepackage{amsmath, amssymb, amsfonts, amsthm, graphicx}
\usepackage[mathscr]{eucal}
\usepackage[symbol, perpage]{footmisc}

@@

\section{The singleton bound and a promising example} \label{s:singleton}
\begin{thm}[Singleton Bound]\label{thm:singleton_bound}
If $C$ is a code with parameters $n, k, d$, then 
	\begin{equation} % IF SPACE NEEDED, TRANSFORM THESE EQUATIONS INTO INLINE
		k + d \le n + 1,
	\end{equation} 
	or equivalently, 
	\begin{equation}
		R + \delta \le 1 + 1/n.
	\end{equation}
\end{thm}

@@

\begin{proof}
	$RS_q(n, k)$ is a subset of $\mathbb{F}_q^n$, thus it has length $n$.  
	Note that $ \mathcal L_k$ is a vector space over $\mathbb{F}_q$.
	Note that $\{1, x, x^2, \dots, x^{k-1}\}$ is a choice of basis for this vector space, thus it has dimension $k$.
	Consider the map $\phi:  \mathcal L_k \to \mathbb{F}_q^n$ given by 
	\begin{equation}
		f \mapsto (f(\alpha_1), f(\alpha_2), \dots, f(\alpha_n)).
	\end{equation}

	Note that the map $\phi$ is a linear transformation.
	Thus, its image $\Ima \phi = RS_q(n, k)$ is also a vector space.  
	Additionally, if $\phi(f) = \phi(g)$, then $f - g$ has at least $n$ roots, but has degree less than $n$; thus $f - g$ is the zero polynomial, which implies $f = g$.
	Therefore, $\phi$ is also injective.
	This implies that the dimension of the domain $ \mathcal L_k$ is the same as the dimension of the image $RS_q(n, k)$, i.e.,  $\dimension RS_q(n, k) = k$.

	Finally, consider distinct $f, g \in  \mathcal L_k$ and define $d = \dist(\phi(f), \phi(g))$, $f - g$ has at least $n - d$ roots.
	Furthermore, $f - g$ is a non-zero polynomial with degree less than $k$, thus has at most $k - 1$ roots.
	Then,  $ k - 1 \ge \#\ \text{roots} \ge n - d$.
	If we choose $f, g$ such that $d$ is the minimal distance, we get $k + d_{\mathrm{min}} \ge n + 1$, which together with the Singleton Bound, implies $k + d_\text{min} = n + 1$.
\end{proof}

@@

\section{Curves, rational functions and divisors}

@@

%\begin{defn}[Rational function]
%	A rational function $f$ over $K$ is the ratio $g/h$ of polynomials $g, h \in K[x, y]$.
%	A homogeneous rational function $F$ over a $K$ is the ratio $G/H$ of homogeneous polynomials $G, H \in K[x, y]$.
%
%	After cancelling common roots of $g, h$, the roots of $g$ are called \emph{zeros} of $f$ and the roots of $h$ are called the \emph{poles} of $f$.  
%	We say $f$ has order $n$ at $P$ if $P$ is a zero of multiplicity $n$; order $-n$ if $P$ is a pole with multiplicity $n$; and order $0$, otherwise.
%\end{defn}

\begin{defn}[Rational function on a curve]
	Let $F \in K[X, Y, Z]$ be a homogeneous polynomial and let $\mathfrak C = \mathfrak C_F$.
	If $G, H \in K[X, Y, Z]$ are homogeneous polynomials of equal degree, then the fraction $G/H$ is called a \emph{rational function} on $\mathfrak C$.
	Fractions $G/H$ and $G'/H'$ define the same rational function if $G'H - GH'$ is a multiple of $F$, i.e., if  $G'H - GH'$ vanishes at all $L$-rational points on $\mathfrak C$.
\end{defn}

	Note that it makes sense to evaluate a rational function on a projective point, since \[
		\frac{G(X_0, Y_0, Z_0)}{H(X_0, Y_0, Z_0)} = \frac{\lambda^d G(X_0, Y_0, Z_0)}{\lambda^d H(X_0, Y_0, Z_0)} = \frac{G(\lambda X_0, \lambda Y_0, \lambda Z_0)}{H(\lambda X_0, \lambda Y_0, \lambda Z_0)},
	\]
	i.e., the result is the same regardless of the representation of a given projective point.

	After cancelling common factors of $G$ and $H$, a $L$-rational point $P$ on $\mathfrak C$ is called a \emph{zero} of $G/H$ whenever $G$ vanishes at $P$, and is called a \emph{poles} of $G/H$ whenever $H$ vanishes at $P$.  
	We say $G/H$ has order $n$ at $P$ if $P$ is a zero of multiplicity $n$; order $-n$ if $P$ is a pole with multiplicity $n$; and order $0$, otherwise.

@@

\begin{defn}[\cite{LS87}]
	If $D = \sum n_p P$, then define $\mathcal L(D)$ as the vector space formed by all homogeneous rational functions $f$ such that the order of $f$ at each $L$-rational point $P$ on $\mathfrak C$ is greater or equal to $-n_P$ (plus the zero polynomial).
\end{defn}

@@

\begin{proof}
	Bezout's theorem implies that $\mathfrak C$ and $\mathfrak C_G$ intersect in the same number of points, counting multiplicity, than $\mathfrak C$ and $\mathfrak C_H$. Therefore, the number of zeros of a function is the same as the number of poles, counting multiplicity. Thus, the sum of the orders of $G/H$ at the $L$-rational points on $\mathfrak C$ must be zero. On the other hand, if $G/H \in \mathcal L(D)$, then the sum of the orders of $G/H$ at the $L$-rational points on $\mathfrak C$ must be $\ge - \deg(D)$.

	Thus, if $\deg(D) < 0$, then a rational function $G/H$ cannot be in $\mathcal L(D)$.
\end{proof}

For our study, an important theorem from Algebraic Geometry that will be useful in Section \ref{s:grs} when creating new types of codes is the following one:
\begin{thm}[Riemann--Roch Theorem, \cite{Wal00}]
	Let $\mathfrak C$ be a nonsingular projective curve of genus%
	\footnote{For the reader that is not familiar with genus, it is enough to know that it is an integer that can be calculated for any given curve.}
	$g$ defined over the field $\mathbb{F}_q$ and let $D$ be a divisor on $\mathfrak C$.
	Then
	\begin{equation}
		\dim \mathcal L(D) \ge \deg D + 1 - g,
	\end{equation}
	with equality holding if $\deg D > 2g - 2$.
\end{thm}

@@

Note that, if a polynomial $f(x) \in \mathcal L_k$ has degree $d < k$, then the rational function \[
	F(X, Y) = \frac{Y^d f(X/Y)}{Y^d}
\]
has $P_\infty$ as its only pole, with order at most $k-1$. Thus, $F(X, Y) \in \mathcal L(D)$ and $f(\alpha_i) = F(P_i)$ for any $\alpha_i$.

@@

\begin{defn}[Algebraic geometric codes, \cite{LS87, Wal00}] %Copied from Wal00
	Let $\mathfrak C$ be an irreducible nonsingular%
	\footnote{These adjectives just make sure $\mathfrak C$ does not have some weird behavior%
	%, e.g., $\mathfrak C$ has no double root%
	. The reader should not worry much about them.}
%%	\todo{Provide an explanation for why the `irreducible nonsingular' is there, but not a thing that the reader should worry. I think the key is some sort of multiplicity.}
	projective plane curve, let $D$ be a divisor on $\mathfrak C$ and let $\mathcal P = \{P_1, P_2, \dots, P_n\}$ be a set of rational points on $\mathfrak C$.
	Assume $\mathcal P$ and $\Supp D$ have no points in common, thus, no $P_i$ can be a pole of any $f \in \mathcal L(D)$, and $f(P_i)$ is well-defined for any $f \in \mathcal L(D)$ and any $P_i \in \mathcal P$.
	Then, the \emph{algebraic geometric code} associated to $\mathfrak C, \mathcal P$ and $D$ is
	\begin{equation}
		C(\mathfrak C, \mathcal P, D) := \{(f(P_1), \dots, f(P_n)) \mid f \in L(D)\} \subset \mathbb{F}_q^n.
	\end{equation}
\end{defn}

@@

	Futhermore, we shall prove that $\psi$ is injective.
	If $\psi(f) = \psi(g)$, then $\psi(f - g) = \vec 0$.
	Let $h := f-g$.
	We have $h(P_1) = h(P_2) = \dots = h(P_n) = 0$, so every $P_i$ is a zero of $h$, and finally $a_{P_i} \ge 1$, for all $1 \le i \le n$.
	Therefore, $h \in \mathcal L(D - P_1 - \cdots - P_n)$.
	Since $\deg D < n$, we have $\deg(D - P_1 - \cdots - P_n) < 0$ and, consequenly, $h$ must be the zero polynomial by Proposition \ref{prop:degDlt0}.
	Therefore, $f = g$, finishing the proof that $\psi$ is injective.
	Since $\psi$ is a injective linear transformation, the dimension of the domain and of the image are the same.
	Therefore, the dimension of $C(\mathfrak C, \mathcal P, D)$ is $k = \dim \mathcal L(D)$, which evaluates to $\deg D + 1 - g$ by the Riemann-Roch Theorem.
